\thispagestyle{empty}

\chapter*{Acknowledgements}

To set off on a thorny path that is not well marked and to be convinced of arriving at destination while discovering lots of intriguing and amazing phenomena, that is called ‘Research’. Pursuing a Ph.D. is a long trip that can be hard and intense if the right people do not accompany us. 

I would like to express my deepest gratitude to my supervisor Abderrahmane Kheddar without whom this journey would not have been so pleasant. His valuable advice, comments, criticisms, and professional leadership during this thesis have considerably improved the quality of this work. His dedication pushed me to try endlessly to meet his expectations and standards, which I hope I did. 

I am very honored to have my thesis reviewed and judged by Jochen Steil, Michael Mistry, Alexander Dietrich, Daniele Pucci and Vincent Padois with their great knowledge and expertise in robotics and control. 

I am also grateful to Dr. Pierre Gergondet and Dr. Arnaud Tanguy for their technical help and assistance concerning the software framework and experiments. The main contributions in this thesis would not have been possible without their effective support.

I wish to thank Dr. Niels Dehio for all the valuable and intense brainstorming discussions that we had at work, at lunch, and even during hiking and canoeing. He often finished our discussions by saying ‘\emph{This is amazing Mohamed!  You should dig more into this}’. His pedagogy always inspired me (even when teaching me how to drive a canoe).  %Thank you very much.

I would also like to extend my sincere thanks to all IDH team colleagues. I had this great chance to meet people from around the world and discover their language and culture.  Saeid Samadi and Yukiko Osawa, my accomplices, for all the funny moments we had during the breaks. You were, and still are, family to me. Yuquan Wang, who initiated me to QP. Alexandra, for your kindness. Antonin Dallard, with whom I shared the office and for all the exciting discussions. Julien Roux, for organizing the team dinners and raclette parties. Ahmed Zermane, for being on my right-hand man. For all the others, I want to say thank you for making the lab-life enjoyable. 

The meeting of wonderful people crowned my life in Montpellier. I am thankful to Bachir and Nadjib for welcoming me on the first days and having oriented me at every moment. Felix and Adrien and their wives for their sense of hospitality and for making me feel one of theirs. 

My special thanks go to my family and their unwavering support. I would like to thank my aunt and uncle for giving me the secret tips for conducting well my research and increasing my willingness. I would also like to thank my brother Walid and his wife for hosting me for the past four years, and without whom the COVID lockdown would not have been easy. Finally, no words will be sufficient to thank my mother for her unconditional love, she who always found the right words to erase my sorrows, and my fiancée, who supported me during all the difficult times. 

   
%\medskip

%\noindent


%\medskip
%\medskip
%\medskip
\medskip

\noindent
\textit{M. D., July 2022, Montpellier, France}